---
title: 'Task 1: Palmetto Binary Logistic Regression (PELEGRI)'
author: "Patrick Pelegri-O'Day"
date: "1/30/2022"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
code_folding: hide
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(GGally)
library(here)
library(cowplot)
library(broom)
library(jtools)
library(caret)
library(kableExtra)
```


A. Overview (done but needs to be reviewed)
B. 3 finalized plots exploring differences in variables for the two species
C. Run binary logistic regression
  - Then perform cross validation
  - Then check AICc
D. Train selected model on entire dataset and show results in a table
E. Show prediction accuracy in a table


### Overview
This report uses binary logistic regression to test the feasibility of using variables plant height, canopy length, canopy width, and number of green leaves to classify whether a palmetto is species Serenoa repens or Sabal etonia. **[Describe the dataset**]

**Data source:** Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

Use binary logistic regression to test the feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia. Use code folding and hide all messages & warnings in your knitted HTML.

### Data exploration

A subset of the data is retained for analysis. The variables considered include species of palmetto, plant height, canopy length, canopy width, and green leaves.
```{r read and wrangle data}
# Read in the data
palmetto_raw <- read_csv(here("data", "palmetto.csv"))

# Take subset for exploration and analysis. Rename species variables by their latin name and convert to factor
palmetto <- palmetto_raw %>% 
  select(species, height, length, width, green_lvs) %>% 
  mutate(species = case_when(
    species == 1 ~ "Serenoa repens",
    species == 2 ~ "Sabal etonia"
  )) %>% 
  mutate(species = as.factor(species)) %>%  # Sabal etonia is 1st, Serenoa repens is 2nd
  drop_na()
  
```

Below, differences in plant height, canopy length, canopy width, and green leaves for the two species, Serenoa repens and Sabal etonia.
CHANGE COLORS

```{r data exploration}
# Explore differences between the two species

explore1 <-ggplot(data = palmetto, aes(x = species, y = green_lvs, fill = species)) + geom_violin(scale = "count", color = "black") + # geom violin to visualize the large volume of data
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) + # adding  median and quartiles to figure
  stat_summary(fun=mean, # adding the mean to the figure 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("cornsilk4", "lightblue3")) + 
  theme_minimal(13) + # change theme and font size
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) + 
  labs(x = element_blank(), y = "Number of green leaves")

explore2 <- ggplot(data = palmetto, aes(x = species, y = height, fill = species)) + geom_violin(scale = "count", color = "black") + 
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) + 
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("cornsilk4", "lightblue3")) + 
  theme_minimal(13) + 
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) + 
  labs(x = element_blank(), y = "Plant height")

explore3 <- ggplot(data = palmetto, aes(x = species, y = length, fill = species)) + geom_violin(scale = "count", color = "black") + 
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("cornsilk4", "lightblue3")) + 
  theme_minimal(13) + 
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) + 
  labs(x = element_blank(), y = "Canopy length")


explore4 <- ggplot(data = palmetto, aes(x = species, y = width, fill = species)) + geom_violin(scale = "count", color = "black") + 
geom_boxplot(color = "black", fill = NA, width = 0.1, outlier.color = NA) + 
  stat_summary(fun=mean, 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  scale_fill_manual(values = c("cornsilk4", "lightblue3")) + 
  theme_minimal(13) + 
  theme(legend.position = "none",
        axis.text.x = element_text(vjust = 5, size = 12)) + 
  labs(x = element_blank(), y = "Canopy width")

plot_grid(explore1, explore2, explore3, explore4, labels = c('A', 'B', 'C', 'D'), label_size = 12, nrow = 2)
```

We see that relationships among palmetto width, height, and length are highly correlated with each other between the two species. However we see that the relationship between palmetto green leaves **[INSERT MORE ACCURATE DESCRIPTION]** and width is distinct between the two palmetto species. Given that green leaves is the only variable that has a different trend between the two species, that is the variable we will focus on for binomial logistic regression.

### Binary logistic regression

```{r binary logistic regression}

# Define two functions that will be the basis of two models
f1 <- species ~ green_lvs + width + height + length
f2 <- species ~ green_lvs + width + height 

# Define two binary logistic regression models based on the functions above
palmetto_blr1 <- glm(formula = f1,
                    data = palmetto,
                    family = 'binomial')

palmetto_blr2 <- glm(formula = f2,
                    data = palmetto,
                    family = 'binomial')

```


```{r cross validation}

set.seed(47)

# Define training method as cross validation, 10 folds, 10 repeats
tr_ctrl <- trainControl(method = 'repeatedcv', number = 10, repeats = 10)

# Train the two models
model1 <- train(f1, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model1

model2 <- train(f2, data = palmetto,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model2

```

Cross-validation results show that Model 1 (accuracy of `r round(model1$results$Accuracy, 1)`) outperforms Model 2 (AICc of `r round(model2$results$Accuracy, 1)`) by approximately ``r round(model1$results$Accuracy - model2$results$Accuracy), 1)` points. Thus, cross-validation indicates that Model 1 is preferable.

```{r AIC}
# Also compare AICc
palmetto_AICc <- AICcmodavg::aictab(list(palmetto_blr1, palmetto_blr2))

palmetto_AICc # AIC results indicate that Model 1 is preferable since it is lower by 807 points
```

AICc results show that Model 1 (AICc of `r round(aic_comparison$AICc[1], 1)`) outperforms Model 2 (AICc of `r round(aic_comparison$AICc[2], 1)`) by approximately `r round(aic_comparison$Delta_AICc[2], 1)` points. This difference is very large - much larger than the significant difference threshold of 2 points. AICc confirms that Model 1 is preferable to Model 2.

```{r train selected model}
# Store tidy results of final model (Model 1)
final_palmetto_blr <- tidy(palmetto_blr1) %>%  
  mutate(p.value = case_when( # rename p-values for readability
    p.value <0.001 ~ "<0.001"
  )) %>% 
  mutate(term = case_when( 
         term == "(Intercept)" ~ "Intercept",
         term == "height" ~ "Height",
         term == "length" ~ "Length",
         term == "width" ~ "Width",
         term == "green_lvs" ~ "# of green leaves")) %>% 
  select(-statistic)

# Create table
kable(final_palmetto_blr, 
      col.names = c("Term", "Coefficient", "Standard Error", "P Value")) %>% 
  kable_minimal(full_width = FALSE)
```

Final model:
`r equatiomatic::extract_eq(final_model_blr, wrap = TRUE, use_coefs = TRUE)`

#### Evaluate model predictions

```{r evaluate model predictions}
# Create function for prediction accuracy
predict_accuracy_f <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  return(accurate)
}
 
# Create fitted outcomes for each model, create column predicted_species based on predictions, create column prediction_result that shows 1 for correct prediction and 0 for incorrect prediction
blr1_predictions <- palmetto_blr1 %>% 
  augment(type.predict = 'response') %>% 
  mutate(predicted_species = case_when(
    .fitted >= 0.5 ~ "Serenoa repens",
    .fitted < 0.5 ~ "Sabal etonia"
  )) %>% 
  mutate(prediction_result = prediction_accuracy(species, predicted_species))

# See what percentage of the time the model predicted correctly
blr1_accuracy <- blr1_predictions %>% 
  group_by(species) %>% 
  summarize(pred_accuracy_v = round(mean(prediction_result), 2))
  
```

**Table 2.** Table caption here
```{r}
# Return results in a table
kable(blr1_accuracy, 
      col.names = c("Species", "% correctly classified")) %>% 
  kable_minimal(full_width = FALSE)
```


```{r}
### PLACEHOLDER… I think I can delete

# Visual comparison of selected model (Model 1): DON'T UNDERSTAND
ggplot(data = blr1_fitted, aes(x = green_lvs, y = .fitted, shape = species)) +
  ### add aes(shape = species) to compare probability with actual
  # geom_point(aes(color = sex, shape = species)) +
  ### add geom_smooth to show general fit
  geom_smooth(se = FALSE) +
  labs(x = "Green leaves",
   	   y = "Probability of outcome Serenoa repens")
```

