---
title: 'Task 2: Parameter Estimation (PELEGRI)'
author: "Patrick Pelegri-O'Day"
date: "1/31/2022"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
code_folding: hide
---

### Overview
[insert overview text: provide purpose and method of the analysis, then describe the dataset.]

For task 2, you will use non linear least squares to estimate parameters of a length to weight model for lizard populations in New Mexico. 

**Data Source:** Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(purrr)
library(tidyverse)
library(Metrics)
library(patchwork)

```

```{r read in data}
lizards_raw <- read_csv(here("data", "lizard.csv"))
```


### Analysis

Our conceptual model is Weight = a(length)$^b$
We want to estimate a and b. To do so, we take the natural log of both sides.
ln(weight) = ln(a(length)$^b$). This is equal to
ln(weight) = b * ln(a) + b * ln(length)

We can use OLS to get an estimate for *a* and *b*.
*b* will simply be the estimated coefficient for ln(length) once we run OLS.
*b* * ln(*a*) will be equal to the intercept estimate from our OLS model.
Thus, the estimate for *a* will be equal to $e^β/b$, where β is the y-intercept estimate

```{r log transform the data}
lizards <- lizards_raw %>% 
  mutate(log_length = log(SV_length)) %>% 
  mutate(log_weight = log(weight))
```

```{r create nls model}

# Create initial OLS model using log variables in order to generate guess of NLS model parameters
my_guess_model <- lm(log_weight ~ log_length, data = lizards)
guess_coeffs <- coef(my_guess_model)

# Define weight to length model for lizards
calc_weight <- function(a, SVL, b){
 W = a*SVL^b
return(W)
}

lizards_nls = nls(weight ~ calc_weight(a, SV_length, b),
                  data = lizards,
                  start = list(a = 2.718^(guess_coeffs[1]/guess_coeffs[2]), 
                               b = guess_coeffs[2]), 
                  trace = TRUE)

kable(lizards_nls,
      caption = "Table 1. NLS model") %>% 
  kable_styling(full_width = FALSE)

```

```{r}
ggplot(lizards_nls)
```

