---
title: 'Task 1: Palmetto Binary Logistic Regression (PELEGRI)'
author: "Patrick Pelegri-O'Day"
date: "1/30/2022"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(GGally)
library(here)
library(cowplot)
```


### Overview
This report uses binary logistic regression to test the feasibility of using variables plant height, canopy length, canopy width, and number of green leaves to classify whether a palmetto is species Serenoa repens or Sabal etonia. **[Describe the dataset**]

**Data source:** Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5

Use binary logistic regression to test the feasibility of using variables plant height (height), canopy length (length), canopy width (width), and number of green leaves (green_lvs) to classify whether a palmetto is species Serenoa repens or Sabal etonia. Use code folding and hide all messages & warnings in your knitted HTML.

### Data exploration

```{r}
# Read in the data
palmetto_raw <- read_csv(here("data", "palmetto.csv"))

# Take subset for exploration and analysis
palmetto <- palmetto_raw %>% 
  select(year, plant, species, height, length, width, green_lvs) %>% 
    mutate(species = as.character(species))
  
```

2 visualizations: explore differences in height, canopy length, canopy width, and green leaves for the two species.

```{r}
# GGpairs (don't include in report)

palmetto %>% 
  select(-year, -plant) %>% 
  ggpairs(aes(color = species))

# Green leaves appear to be the only differentiator between the species
```


```{r}
# Explore differences between the two species

explore1 <- ggplot(palmetto, aes(x = width, y = green_lvs, color = species)) +
  geom_point() +
  theme_minimal()

explore2 <- ggplot(palmetto, aes(x = height, y = length, color = species)) +
  geom_point() + 
  theme_minimal()

explore3 <- ggplot(palmetto, aes(x = width, y = height, color = species)) +
  geom_point() +
  theme_minimal()

plot_grid(explore1, explore2, explore3, labels = c('A', 'B', 'C'), label_size = 12, nrow = 2)
```

We see that relationships among palmetto width, height, and length are highly correlated with each other between the two species. However we see that the relationship between palmetto green leaves **[INSERT MORE ACCURATE DESCRIPTION]** and width is distinct between the two palmetto species.
